A programming model for temporal coordination (in music)

### 2014-05-06 paper/paper.tex
Add Alan's intro and associated refs

@@

%
\lang{} was previously discussed in~\cite{aaron2013sonic}

>>>>

Sonic Pi is a music live coding language that has been designed for
educational use, as a first programming language~\cite{Aaron2013}. A first
programming language should be conceptually straightforward and
syntactically uncluttered. However, it is not straightforward to achieve
this simplicity in a music live coding language, for reasons largely
related to the representation of time. Representing musical time in a
programming language is often problematic, firstly because natural ways
of describing and organising musical events are multi-threaded (scores,
chords, resonance, reverb), and secondly because so many standard
computational formalisms treat execution time as a non-functional
requirement. Time can be even more problematic in live coding, because
the creation of the code, as a performance, is interleaved with the
sound events resulting from its execution. Yet for users of Sonic Pi,
the creative experience they have, like all experience, arises through
time - as media theorist Mieke Bal says, "time is where subjectivity is
produced"~\cite{Bal2002}.

As noted by Julian Rohrhuber~\cite{blackwell_et_al2014}, there have been
many publications discussing alternative approaches to the
representation of time in live coding, many having to choose between
explicit or implicit representation of time, and between description of
time with reference to internal or external state. These many
interesting strategies include McLean's formalism of cyclic time in the
Tidal language~\cite{McLean2013}, and Sorensen's temporal recursion in
Impromptu/Extempore~\cite{Sorensen2010}. In this paper, we present a formalism that is
designed to achieve production-quality sound (via the SuperCollider
synthesis server) while allowing inexperienced programmers in an
educational setting (typically 11-12 year-old children) to express the
temporal structure in terms that have an intuitive correspondence to the
experience and production of musical sounds.

@@

%
\lang{} was previously discussed in~\cite{Aaron2013}
